### 2011-06-12 thesis/novelty.tex
Some cleanup

\subsection{Old content}

On this section it will be presented how to perform novelty detection of a
given variable class assuming samples are drawn from a distribution described
by the graph.

On the previous section it was shown how to detect a novel class on a variable
assuming that the rest of the graph variables and structure is correctly modeled
by the agent knowledge.
In this section methods are presented that not only allow to extended this to
detection on multiple variables but also allow the system to incorporate
novelty information back in the graphical model in order to continue its
probabilistic and uncertain reasoning.

When trying to extend the method to deal with multiple novel variables it may
be tempting to follow a deterministic approach such as:
when the agent determines a variable $a$ to have a novel class, it marks it as
novel and proceeds from that moment by using a graphical model where the factors
associated with the $a$ replaced by unconditional factors.
But such an approach would not be able to optimally detect novel variables
as later decisions are dependent on former ones.
It would also not be able to adapt itself and reconsider previously decisions
when more data becomes available for the agent.

To handle those issues a fully probabilistic approach should be taken
instead of a deterministic.
That can be done by extending all the variables on the factor graph representing
the semantic map with an additional state \emph{novel} and extending all the
factors with information on the unconditional distribution.

Such a factor graph would be able to handle novel variables and incorporate
prior-information on the unconditional distributions since it is able to
represent any state of the $X$ variables present on the graph where any
subset of them is novel. But in order to add the ability to control the
novelty thresholds a normalization factor $\phi_N$ needs to be introduced.

\begin{figure}[h]
\centering
Yeah get a figure for here.
\caption{\label{fig:multiple-big-threshold}Factor graph extended to be able to
         represent novelty on any subset of variables. Note the added factor
         $\phi_N$ that servers as a normalization factor such that probability
         with any subset of $Y$ variables considered novel can be scaled with
         any value such that any thresholding decision mechanism can be
         implemented.}
\end{figure}

In order to give full control on the threshold on which any subset of
variables is considered novel the factor $\phi_N$ needs to be dependent on all
the $X$ variables and act as a weighting function that correctly scales the
likelihood of a given instance of the $X$ variables.
If the criterium for defining the threshold is the likelihood of all variables
$X$ being non-novel has the max likelihood if and only if
\autoref{eq:multiple-threshold} holds for all the non-empty subsets $Y$ of $X$.
The weight for $\phi_N$ when the set $Y$ is classified as novel
can be set for $t_Y$ and $\phi_N$ for the empty subset of novel variables is set
to $1$.

\begin{equation}
\label{eq:multiple-threshold}
\frac{P_G(x)}{P_{U_Y}(x)} > t_Y
\end{equation}

Still without further assumptions a threshold $t_Y$ would need to be defined for
all the $2^N-1$ non-empty subsets.
By taking a simple approach of considering the novelty on a variable independent
from the novelty on any other variable $t_Y$ can be factorized:

\begin{equation}
t_Y = \prod_{i \in Y}{t_i}
\end{equation}

This factorization yields then the interesting property to allow to split the
threshold factor $\phi_N$ into $N$ different $\phi_{N_i}$ associated with each
variable of the graph. A graph able to handle novelty detection on several
variables with the described factorization is illustrated on
\autoref{fig:multi-factorized-threshold}.

\begin{figure}[h]
\centering
Yeah get a figure for here.
\caption{\label{fig:multi-factorized-threshold}Factor graph extended to be able
         to represent novelty on any subset of variables. Note the added factor
         $\phi_N$ has been split into a $\phi_{N_i}$ still it is able to perform
         detection on multiple variables according to the factorization
         described in the text.}
\end{figure}

To sum up, the method described above allows to integrate novelty detection
into the graph, providing the agent with tools to describe its belief on the
novelty of variables on the system and to continue working with such a
representation. The mechanism is also fully probabilistic and at no time
the agent needs to take an irrevocable decision on the novelty of a variable,
allowing the agent beliefs to change over time as more data becomes available.

@@

\section{Detecting New Structure}

>>>>

\subsection{Old content}
To sum up, the method described above allows to integrate novelty detection
into the graph, providing the agent with tools to describe its belief on the
novelty of variables on the system and to continue working with such a
representation. The mechanism is also fully probabilistic and at no time
the agent needs to take an irrevocable decision on the novelty of a variable,
allowing the agent beliefs to change over time as more data becomes available.

@@

\section{Detecting New Structure}
I do not belive there is any time left for this.
